---
title: "ESM 206 Lab 10"
author: "Anna PW"
date: "December 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

# PART 0. Load libraries
```{r packages}

library(tidyverse) # The tidyverse!
library(corrplot) # For correlation matrices
library(janitor) # For cleaning up column names
library(lubridate) # For dealing with dates & times
library(gganimate) # For adding animation to ggplot graphics
library(ggimage) # For updating graph components with images


```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr}
en_var <- read.csv("world_env_vars.csv") %>%
  na.omit
#View(en_var)

#Calculate pearson's R for correlations. 
cor_df <- cor(en_var[2:28]) #correlation variables for columns 2-28
cor_df

#plot the correlation variables/correlation matrix
corrplot(cor_df, 
         type = "upper", 
         method = "ellipse", 
         tl.col = "black", 
         tl.cex = 0.5) #change text label color and size

#Reading the chart: 
#narrower the ellipse and darker the color, the stronger the correlation
#blue is pos correlation
#red is neg correlation
```

# PART 2. Binary Logistic Regression (Donner Party Data)

Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$

We'll use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done. 

a. Read in the DonnerTable.csv file as DonnerTable

```{r donner_data}
donnertable <- read.csv("DonnerTable.csv")
View(donnertable)
# 0 death
# 1 survival

# specify binary variables using "casewhen" to create a new column for 0 and 1 outcomes

```

b. Binomial logistic regression

```{r blr}

donnerblr <-glm(Outcome ~ Sex + Age, family = "binomial", data = donnertable)
summary(donnerblr)
```

c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 
#simplified dataframe 
f5 <- data.frame(Sex = "Female", Age = 5)


# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":

f5logodds <- predict(donnerblr, newdata = f5, type = "link") #link calcs log odds

# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

f5odds <- exp(f5logodds)

#80% chance of survival odds are 1 to 4.23

# Ask: Could we manually find the probability of survival for a 5-year old female? recall: p/(1-p) = ODDS

# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:

f5prob <- predict(donnerblr, newdata= f5, type = "response")
f5prob
```

d. What is the probability of survival for a 25 year-old male? 

```{r predict_m_25}

# Similarly:

m_25 <- data.frame(Sex = "Male", Age = 25) # Make a new data frame

m_25_prob <- predict(donnerblr, newdata = m_25, type = "response") # Find probability of survival
m_25_prob

```

e. Create new sequences of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a sequence from 0 to 100, twice (one will be "Male" and one will be "Female")

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex")
View(donner_newdata)
```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE.

predictedprobs <- predict(donnerblr, newdata = donner_newdata, type = "response", se.fit = TRUE)

# Coerce outcome into data frame. 

graphdata <- data.frame(donner_newdata, predictedprobs$fit, predictedprobs$se.fit)
colnames(graphdata) <- c("Age", "Sex", "Probability", "SE")
View(graphdata)
```

g. Graph results.

```{r donner_graph}
ggplot(graphdata, aes(x= Age, y = Probability))+
  geom_line(aes(color = Sex))+ #separates male and female by line and color
  geom_ribbon(aes(ymin = Probability - SE, ymax = Probability + SE, fill = Sex), alpha = 0.5)+
  theme_classic()
  
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}

sifull <- list.files(pattern = "solar_irradiation_*")%>%#pulls files starting with "solar_irr..."
  map_df(~read_csv(.)) %>%#use the outcome of the previous line "."
  clean_names() #simplifies column names



```

b. Wrangle the data
```{r wrangle_solar}
solar_tidy <- sifull %>%
  rename(solrad = etr_wh_m_2, 
         date = yyyy_mm_dd, 
         time = hh_mm_lst) %>% #rename columns
  filter(time != "NA") %>% #remove NA values
  mutate(site = fct_relevel(site, "Hawaii", "Santa Barbara", "Alaska")) #change order of sites 


```

c. Use lubridate() functions to convert to times/dates
```{r date_time}
solar_tidy$date <- mdy(solar_tidy$date) #assigns date column so R recognizes it as dates 

```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r solar_graph}

solargg <- ggplot(solar_tidy, aes(x= date, y = time)) +
  geom_tile(aes(fill= solrad)) +
  scale_fill_gradientn(colors = c("royalblue2", "mediumorchid1", "orange", "yellow")) + #allows you to specify colors
  scale_y_time() +
  facet_wrap(~site, ncol = 1)
solargg

```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States, Brazil, Chile, and Ecuador

a. Get the data, and do some wrangling:
```{r aqua_data}
aqdf <- read_csv("aq_wb.csv")
#View(aqdf) #not in tidy format b/c years are in x axis, it's wide format

aqtidy <- aqdf %>%
  filter(country == "Brazil" | #select specific country
           country == "Chile" |
           country == "Ecuador" |
           country == "United States") %>%
  gather(year, aq_prod, `1960`:`2016`) %>%
  filter(year >= 1990) %>%
  mutate(aq_mil =aq_prod/1000000) %>%
  select(country, year, aq_mil)

```


b. Read in the fish.png as 'fish'

```{r fish_image}

fish <- "fish.png"
```

c. Make a graph...with gganimate!
```{r gganimate_fish}
aqplot <- ggplot(aqtidy, aes(x= as.numeric(year), y= aq_mil, group = country)) +
  geom_line(aes(color=country)) +
  geom_image(aes(image = fish)) +
  geom_text(aes(label = country, color= country), position = position_nudge(y=0.04, x=1), size = 5) +
  transition_reveal(country, as.numeric(year))
aqplot

```

##END LAB